---
title: "Market Signals"
author: "Antony"
date: "2021-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import}
suppressPackageStartupMessages({
  import(rpkgs)
})
import(db)
```

# Deriving features for "market effects"

The target variable in this competition is

$$
R(t) = \texttt{ln}\left(\frac{P(t+16)}{P(t+1)} - M(t)\right)
$$

where

* $R(t)$ return of asset at $t$
* $P(t)$ price of asset at $t$
* $M(t)$ return of market at $t+16$

Put simply, it's "(return of the asset) minus (return of the market)".

Without any features for this, the model would have to infer the market state
from the features for a particular asset, makes its job harder.

Instead, lets help the model out by providing some information about the whole market.

# Features

```{r feature}
windowSize = 42 # minutes
df = getQuery(glue('
WITH
t1 AS (
  SELECT
    ts
  , close - AVG(close)
    OVER (PARTITION BY asset_id ORDER BY ts ASC ROWS {windowSize} PRECEDING)
    AS close_centred
  , STDDEV_SAMP(close)
    OVER (PARTITION BY asset_id ORDER BY ts ASC ROWS {windowSize} PRECEDING)
    AS close_stddev
  FROM trn
),
t2 AS (
  SELECT
    *
  , close_centred / NULLIF(close_stddev, 0)
    AS close_std
  FROM t1
),
t3 AS (
  SELECT
    ts
  , AVG(close_std) AS market_close_std
  , AVG(close_stddev) AS market_close_stddev
  FROM t2
  GROUP BY ts
),
t4 AS (
  SELECT
    *
  , REGR_SLOPE(market_close_stddev, EXTRACT(EPOCH FROM ts)::REAL)
    OVER (ORDER BY ts ASC ROWS {windowSize} PRECEDING)
    AS market_close_stddev_slope
  , LEAD(market_close_std, {windowSize})
    OVER (ORDER BY ts ASC)
    AS lead_market_close_std
  FROM t3
)
SELECT * FROM t4
'))

df
```

* `market_close_std` standardised close price over last `r windowSize` minutes, averaged across assets
* `market_close_stddev` standard deviation of close price over last `r windowSize` minutes, averaged across assets
* `market_close_stddev_slope` slope of `market_close_stddev` over last `r windowSize` minutes

# Time series plot

```{r time plots}
set.seed(25582)
viewSize = 200
randomDateIdx = sample(viewSize:nrow(df), 1)
randomDateIdxs = (randomDateIdx - viewSize):randomDateIdx
p =
  df[randomDateIdxs,] |>
  melt(id.vars = "ts") |>
  ggplot(aes(ts, value)) +
  geom_line() +
  facet_wrap(~variable, ncol = 1, scales = "free")

ggplotly(p)
```

# Correlation plots

Compare the correlation against `market_close_std` `r windowSize` minutes in the future.

```{r corr plots}
pp = list()
for (f in c("market_close_std", "market_close_stddev", "market_close_stddev_slope")) {
  featureCorrelation = cor(df[,get(f)], df[,lead_market_close_std], use = "complete.obs")
  p =
    df |>
    ggplot(aes_string(f, "lead_market_close_std")) +
    geom_bin_2d() +
    labs(
      title = glue('{f} -> lead_market_close_std (corr: {featureCorrelation})')
    )
  
  pp[[f]] = p
}
```

```{r}
ggplotly(pp[["market_close_std"]])
```

```{r}
ggplotly(pp[["market_close_stddev"]])
```

```{r}
ggplotly(pp[["market_close_stddev_slope"]])
```

The linear correlations are not very strong, will be interesting to see what the model does with it.
