---
title: "Model: Baseline: target = 0 (mae: 0.00177)"
author: "Antony"
date: "2021-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r import}
suppressPackageStartupMessages({
  import(rpkgs)
})

import(run)
```

Baseline "model" just predict the target is always zero.

Note about the target, recall that the target is in log-space, so target = zero
the same as saying "forecast that price will not change".

# Model

```{r model}
modelName = "baseline-zero"

assets = getAllAssets()

runModel = \() {
  doRun(
    name = modelName,
    trnAmt = 0, # no training data, we always forecast 0
    tstAmt = 60 * 24 * 7 * 2, # 2 weeks, submission period will provide new data every 2 weeks
    assets = assets[,asset_id],

    makeData = \(env, minDate, maxDate, assets, ...) {
      selectStmt = glue('
        SELECT ts, asset_name, target
        FROM trn
        WHERE (ts BETWEEN $1 AND $2)
          AND asset_id IN ({paste(assets, collapse = ", ")})
      ')

      df = getQuery(selectStmt, params = list(minDate, maxDate))
      
      # these are passed through as `trn` and `tst` for training and
      # inference, respectively
      # NB: in this case we don't actually use `x`, but doRun requires that we
      # set something
      env$x = df[,.(ts, asset_name)]
      env$y = df[,.(target)]
    },

    trainModel = \(model, trn, ...) {
      # normally would train a model in here, but we always forecast zero,
      # so nothing to do
      
      # give the model a description
      model$description = 'target(asset, t) = 0'
    },

    predictModel = \(model, tst, ...) {
      # use advanced machine learning algorithm to predict crypto movement
      tst$yhat = vector(mode = "numeric", length = nrow(tst$x))
      tst$yhat[1:length(tst$yhat)] <- 0
    }
  )
}
```

# Method

I have defined a function called `doRun` which picks a random date inside the 
competition data (`t1`), trains a model using the data between `(t1 - trnAmt):t1`,
then evaluates the model using the data between `(t1 + 15):(t1 + 15 + tstAmt)`.
(The ` + 15` is because we are doing 15-minute forecasts, avoid leakage).

You can check out the code in GitHub if you are interested, [link here](https://github.com/KyussCaesar/kaggle-cryptoforecast/blob/master/run.R)

I will run this function a number of times in order to get a distribution of the
evaluation metrics.

```{r method, results=FALSE, warning=FALSE, cache=TRUE}
numSamples = 610

set.seed(205794)

for (i in 1:numSamples) {
  results = runModel()
}
```

# Plots

We can examine the results from the last run, as a sanity-check.

```{r sanity, results=FALSE, cache=TRUE}
df = results$tst$x
df$y = results$tst$y$target
df$yhat = results$tst$yhat
DT::datatable(df[sample(200)])

# sample of data
plotStart = sample(df[,ts], 1)
plotEnd = plotStart + as.difftime(200, units = "mins")

assets[sample(nrow(assets), 2),asset_name] |>
  lapply(\(asset) {
    df[asset_name == asset & ts > plotStart & ts < plotEnd] |>
      melt(id.vars = c("ts", "asset_name")) |>
      ggplot(aes(ts, value, colour = variable)) +
      geom_line() +
      facet_wrap(~asset_name, ncol = 1)
  }) |>
  print()
```

Yep, they all zeros.

# Evaluation

`doRun` writes results to the DB, we can pull them from there.

```{r}
scores = getQuery('SELECT * FROM metrics WHERE name = $1', params = list(modelName))
DT::datatable(scores[,.(run_id, corr, mae, aae, rmse)])
```

Hmm, notice the correlation is missing for all of them? Is there a bug?
No; you cannot compute the correlation between two variables if one of them has
zero variance. In our case we always predict zero, the variance in our predictions
is zero and therefore the correlation is undefined.

Probably that's "um, no shit sherlock" to stats people but it was news to me haha.

Let's take a look at Median Absolute Error instead, which I have abbreviated as `mae`
just to be confusing.

```{r}
p = ggplot(scores, aes(mae)) +
  geom_histogram()
ggplotly(p)
```

MAE is pretty small already, gonna be a tough competition!!

Mean mae: `r mean(scores[,mae], na.rm = TRUE)`
